\documentclass{segabs}

% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

\begin{document}

\title{An example SEG expanded abstract}

\renewcommand{\thefootnote}{\fnsymbol{footnote}} 

\author{Joe Dellinger\footnotemark[1], BP 
  and Sergey Fomel, University of Texas at Austin}

\footer{Example}
\lefthead{Dellinger \& Fomel}
\righthead{SEG abstract example}

\maketitle

\begin{abstract}
  This is an example of using \textsf{segabs.cls} for writing
  SEG expanded abstracts.
\end{abstract}

\section{Introduction}

This is an introduction. \LaTeX\ is a powerful document typesetting
system \cite[]{lamport}. An excellent reference is \cite[]{kopka}. The
new \textsf{segabs.cls} class complies with the \LaTeX2e\ standard.

\section*{Data and Methods}

The data mostly come from portable broad-band stations of the Brazilian Lithosphere Seismic Project, deployed across several parts of Brazil since 1992. and publicly available from the IRIS (Incorporated Research Institutions for Seismology) data management system. A few permanent stations of the Global Seismographic Network and the GEOSCOPE network are also included (Fig. 1). Specifically, our data set contains 53 stations from which we select the vertical component, thus focusing the study on Rayleigh waves. The seismic records span a 16-yr interval, from 1996 to 2012. As Fig. 1 shows, the stations most densely cover central and southeastern Brazil, that is, the São Francisco craton, the adjacent Tocantins and Mantiqueira provinces
and the Paraná basin. This being said, stations located further to the north are crucial to provide coverage of the Amazon craton, the Parnaı́ba basin and the Borborema province.


\subsection{Cross-correlation}

As mentioned in the Introduction, the cross-correlation function
ρ AB (τ ) between the ambient noise u A (t), u B (t) recorded at two seis-
mic stations located at A and B,

Here are examples of equations involving vectors and tensors:
\begin{equation}
\tensor{R} = 
\pmatrix{R_{\rs{XX}} & R_{\rs{YX}} \cr R_{\rs{XY}} & R_{\rs{YY}}} 
=
\tensor{P}_{M\rightarrow R} \; \tensor{D} \; \tensor{P}_{S\rightarrow M}
\;\;\; \tensor{S} \ \ \  ,
\label{SVD}
\end{equation}
and
\begin{equation}
R_{j,m}(\omega) =
\sum_{n=1}^{N} \, \,
P_{j}^{(n)}(\mathbf{x}_R) \, \,
D^{(n)}(\omega) \, \,
P_{m}^{(n)}(\mathbf{x}_S) \ \ \ .
\label{SVDray}
\end{equation}

tends to the Green’s function between A and B (Shapiro \& Campillo 2004): the positive lag part of ρ AB (τ ) gives the response at B to an impulse located at A, and vice versa for the negative lag part. In practice, several signal processing steps are required before the cross-correlation between each pair of stations can be estimated. We follow in the present work the steps advocated by Bensen et al. (2007), which consist in stacking daily cross-correlations after (1) removing the instrument response, the mean and the trend, (2) band- pass filtering the data between 7–60 or 3–60 s, (3) downsamplingthe signal to 1 Hz, (4) normalizing the signal by the running average of its absolute values and (5) whitening the Fourier amplitude
spectrum in the considered frequency band.

In step 2, we found that periods up to at least 60 s must be retained in the waveforms in order to maximize the signal-to-noise ratios (SNRs) of the cross-correlations up to periods 20–25 s. Furthermore, the period band 3–60 s serves to produce the group velocity map at the shortest periods, 6 and 10 s (see Section 2.4), but we observed that the inclusion of high frequencies yields spectralleakage effects that significantly degrade the SNRs at periods larger than 10 s. The tomographic inversions at longer periods are therefore based on the data bandpass filtered between 7 and 60 s. Further
work is needed to identify the causes of the spectral leakage, which we suspect to be related to the time-domain normalization (step 4).This step is intended to minimize the influence of phenomenons that obscure ambient noise, such as earthquakes and instrumental irregularities. Bensen et al. (2007) note that the normalization weights may be computed from the waveform bandpass filtered in the earthquake period band in order to better remove earthquakes close to the noise level. As the level of seismicity is not significant in the studied region, this did not yield any improvement on our data (and even slightly degraded the SNRs at short periods), and we thus normalized the signal using the entire considered period band. Many studies apply the ‘one-bit normalization’ technique, wherein only the sign of the signal is retained (+1 or −1) and the amplitude is disregarded completely (e.g. Shapiro \& Campillo 2004; Picozzi et al. 2009; Huang et al. 2010). We found that the running-absolute-mean normalization yields much higher SNRs than the one-bit normalization on our data set, which is in agreement with the results of Seats et al. (2012).

Fig. 2 shows all the available cross-correlations, normalized to unity and stacked as a function of the inter-station distance. Arrival times corresponding to propagation speeds between 2.5 and 3.5 km s −1 clearly emerge from the cross-correlations. As a last step before measuring the dispersion curves, the cross-correlations are symmetrized and turned into one-sided signals by averaging the positive and the negative lag parts. This symmetrization method was applied in most previous studies, but alternative techniques could be tested in future research: for example, discarding of the
side with the lowest SNR (Pawlak et al. 2011) or phase coherence weighting (Schimmel \& Gallart 2007; Dias et al. 2015).

\subsection*{Figures}
\renewcommand{\figdir}{Fig} % figure directory

Figure~\ref{fig:waves} shows what it is about.

\plot{waves}{width=\columnwidth}
{This figure is specified in the document by \texttt{
    $\backslash$plot\{waves\}\{width=$\backslash$columnwidth\}\{This caption.\}}.
}

\subsubsection{Multiplot} 

Sometimes it is convenient to put two or
more figures from different files in an array (see
Figure~\ref{fig:exph,exgr}). Individual plots are
Figures~\ref{fig:exph} and~\ref{fig:exgr}.

\multiplot{2}{exph,exgr}{width=0.4\textwidth}
{This figure is specified in the document by \texttt{
    $\backslash$multiplot\{2\}\{exph,exgr\}\{width=0.4$\backslash$textwidth\}\{This caption.\}}.
}

The first argument of the \texttt{multiplot} command specifies the
number of plots per row.

\subsection{Tables}

The discussion is summarized in Table~\ref{tbl:example}.

\tabl{example}{This table is specified in the document by \texttt{
    $\backslash$tabl\{example\}\{This caption.\}\{\ldots\}}.
}{
  \begin{center}
    \begin{tabular}{|c|c|c|}
      \hline
      \multicolumn{3}{|c|}{Table Example} \\
      \hline
      migration\rule[-2ex]{0ex}{5ex} & 
      $\omega \rightarrow k_z$ & 
      $k_y^2+k-z^2\cos^2 \psi=4\omega^2/v^2$ \\
      \hline
      \parbox{1in}{zero-offset\\diffraction}\rule[-4ex]{0ex}{8ex} &
      $k_z\rightarrow\omega_0$ &
      $k_y^2+k_z^2=4\omega_0^2/v^2$ \\
      \hline
      DMO+NMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_0$ & 
      $\frac{1}{4}
      v^2k_y^2\sin^2\psi+\omega_0^2\cos^2\psi=\omega^2$ \\
      \hline
      radial DMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_s$ &
      $\frac{1}{4}v^2k_y^2\sin^2\psi+\omega_s^2=\omega^2$\\
      \hline
      radial NMO\rule[-2ex]{0in}{5ex} & $\omega_s\rightarrow\omega_0$ &
      $\omega_0\cos\psi=\omega_s$\\
      \hline
    \end{tabular}
  \end{center}
}

\section{ACKNOWLEDGMENTS}

I wish to thank Ivan P\v{s}en\v{c}\'{\i}k and Fr\'ed\'eric Billette
for having names with non-English letters in them.  I wish to thank
\cite{Cerveny} for providing an example of how to make a bib file that
includes an author whose name begins with a non-English character and
\cite{forgues96} for providing both an example of referencing a Ph.D.
thesis and yet more non-English characters.

\append{Appendix example}

According to the new SEG standard, appendices come before references.

\begin{equation}
\frac{\partial U}{\partial z} = 
\left\{
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial g}\right]^2} +
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial s}\right]^2}
\right\}
\frac{\partial U}{\partial t}
\label{eqn:partial}
\end{equation}
It is important to get equation~\ref{eqn:partial} right.

\append{Another appendix}

\begin{equation}
\frac{\partial U}{\partial z} = 
\left\{
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial g}\right]^2} +
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial s}\right]^2}
\right\}
\frac{\partial U}{\partial t}
\label{eqn:partial2}
\end{equation}
Too lazy to type a different equation but note the numeration.

The error comparison is provided in Figure~\ref{fig:errgrp}.

\plot*{errgrp}{width=0.8\textwidth} {This figure is specified in the
  document by \texttt{
    $\backslash$plot*\{errgrp\}\{width=0.8$\backslash$textwidth\}\{This caption.\}}.  }

\onecolumn

\append{The source of the bibliography}
\verbatiminput{example.bib}

\twocolumn

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{example}

\end{document}
